---
title: "p8105_hw2_rh3064"
author: "Rahul Hosalli"
date: "`r Sys.Date()`"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

# Problem 2

The Mr. Trash Wheel data set is imported using `read_excel`. As the excel file contains multiple sheets of data, the first sheet is read, and a range is provided to only include data and omit figures/graphics, empty cells and the final row, which is not dumpster-specific. This is piped to `janitor::clean_names()` .

The `janitor::clean_names()` function is used to convert the column names into snake_case for ease of use later. This is piped to `mutate()`.

The `mutate()` function is used to modify the sports_balls column. In this case rounding to the nearest whole numbers and then changing the variable class to integer. It is also used to create a new column, "trash_wheel", which will keep track of the original data source when combined with the professor trash wheel sheet. Finally, the class of year is changed to double from character.

```{r MrTrashWheel_Import}
mr_trash_wheel<-
  read_excel("./Data/Trash Wheel Collection Data.xlsx", 1, range = "A2:N549") %>%
  janitor::clean_names() %>%
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    
    trash_wheel = c("mr"),
    
    year = as.double(year)
  )
head(mr_trash_wheel)
```

This same process is repeated below for professor trash wheel.

```{r ProfTrashWheel_import}

prof_trash_wheel <-
  read_excel("./Data/Trash Wheel Collection Data.xlsx", 2, range = "A2:M96") %>%
  janitor::clean_names() %>%
  mutate(
    trash_wheel = c("prof")
  )
head(prof_trash_wheel)
```

Finally the data sets are combined using `bind_rows()`. Summary statistics needed are calculated using a series of piped functions to `summarise()` .

```{r Combining_trashWheel_data}
trash_wheel_tidy <- bind_rows(mr_trash_wheel, prof_trash_wheel)

trash_wheel_tidy %>% 
 group_by(trash_wheel) %>%
  summarise(
    weight_sum = sum(weight_tons),
  )

trash_wheel_tidy %>%
  filter(year == 2020) %>%
  group_by(trash_wheel) %>%
  summarise(sports_balls_sum = sum(sports_balls))



```

The data contains `r nrow(trash_wheel_tidy)` observations and `r ncol(trash_wheel_tidy)` variables. Key variables include *dumpster*, which lists the dumpster number, *year* and *date* variables, *weight_tons* (weight collected in tons)and *volume_cubic_yards*. There are also variables the break down the type of trash collected at each instance.

The total weight of trash collected by Professor Trash Wheel is 856 tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.
