---
title: "p8105_hw2_rh3064"
author: "Rahul Hosalli"
date: "`r Sys.Date()`"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

# Problem 2

The Mr. Trash Wheel data set is imported using `read_excel`. As the excel file contains multiple sheets of data, the first sheet is read, and a range is provided to only include data and omit figures/graphics, empty cells and the final row, which is not dumpster-specific. This is piped to `janitor::clean_names()` .

The `janitor::clean_names()` function is used to convert the column names into snake_case for ease of use later. This is piped to `mutate()`.

The `mutate()` function is used to modify the sports_balls column. In this case rounding to the nearest whole numbers and then changing the variable class to integer. It is also used to create a new column, "trash_wheel", which will keep track of the original data source when combined with the professor trash wheel sheet. Finally, the class of year is changed to double from character.

```{r MrTrashWheel_Import}
mr_trash_wheel<-
  read_excel("./Data/Trash Wheel Collection Data.xlsx", 1, range = "A2:N549") %>%
  janitor::clean_names() %>%
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    
    trash_wheel = c("mr"),
    
    year = as.double(year)
  )
head(mr_trash_wheel)
```

This same process is repeated below for professor trash wheel.

```{r ProfTrashWheel_import}

prof_trash_wheel <-
  read_excel("./Data/Trash Wheel Collection Data.xlsx", 2, range = "A2:M96") %>%
  janitor::clean_names() %>%
  mutate(
    trash_wheel = c("prof")
  )
head(prof_trash_wheel)
```

Finally the data sets are combined using `bind_rows()`. Summary statistics needed are calculated using a series of piped functions to `summarise()` .

```{r Combining_trashWheel_data}
trash_wheel_tidy <- bind_rows(mr_trash_wheel, prof_trash_wheel)

trash_wheel_tidy %>% 
 group_by(trash_wheel) %>%
  summarise(
    weight_sum = sum(weight_tons)
  )

trash_wheel_tidy %>%
  filter(year == 2020) %>%
  group_by(trash_wheel) %>%
  summarise(sports_balls_sum = sum(sports_balls))



```

The data contains `r nrow(trash_wheel_tidy)` observations and `r ncol(trash_wheel_tidy)` variables. Key variables include *dumpster*, which lists the dumpster number, *year* and *date* variables, *weight_tons* (weight collected in tons)and *volume_cubic_yards*. There are also variables the break down the type of trash collected at each instance.

The total weight of trash collected by Professor Trash Wheel is 190 tons. The total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.

# Problem 3

The pols-month data is read in via `read_csv`. `separate()` is used to break the *mon* variable in *year*, *month* and *day*. *month.abb* is a built-in constant that contains the English abbreviations for the months of the year, starting from January. The *month* variable is first coerced into a numeric vector, allowing it to be used inside the brackets to refer to the corresponding month name (e.g. if the month was 3, it would refer to the third element of *month.name* and be reassigned to 'Mar').

A new *president* variable is created inside a `mutate()` statement using `case_when()`. The logic inside `case_when()` is only reliant on the *prez_dem* variable as the *prez_gop* variable has a range of values from 0-2 (Ford took office in 1974 after Nixon stepped down, resulting in a value of 2 for Aug-Dec 1974).

Finally, *prez_dem* *prez_gop* and *day* are removed via `select()`.

```{r}
pols_month <- read_csv("./Data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate (
    month = month.abb[as.numeric(month)],
    president = case_when(prez_dem == 1 ~ "dem",
                          prez_dem == 0 ~"gop"),
    year = as.double(year)
    ) %>%
  select(-prez_dem, -prez_gop, -day)

```

Next, the snp data set is imported and tidied in a similar manner as above. `relocate()` is used to reorder the key variables. In this case, the date variable is formatted as MM/DD/YY, so after separation *year* will only be a two digit value. In order to correct this, `case_when()` is used to add 2000 to the *year* if it is \< 50, or 1900 if it is less ≥ 50. The data is then ordered properly for merging with `relocate()`.

```{r}
snp <- read_csv("./Data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into= c("month", "day", "year"), sep = "/") %>%
  mutate(
    month = month.abb[as.numeric(month)],
    year = as.double(year),
    
    year = case_when(year < 50 ~ 2000 + year,
                      year >= 50 ~ 1900 + year),
    ) %>%
  select(-day) %>%
  relocate(year, month)
```

Next, the unemployment data is imported and tidied. The data is in a white format, so `pivot_longer()` is used to to turn it into tidy data.

```{r}
unemployment <- read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  mutate(year = Year) %>%
  select(year, month, unemployment)


  
```

Finally the data sets are joined sequentially. `left_join()` is used in each instance. snp is joined into pols-month, and then unemployment is joined into that dataset.

```{r}
p3_j1 <- left_join(pols_month, snp)
p3_final <- left_join(p3_j1, unemployment)
```
